---
title: "R Notebook"
output: html_notebook
---


## Players covered with tattoos

```{r}
pct_body <- c(10.8, 14.1, 17.6, 19.3, 15.4, 15.3, 15, 17.8, 13.5, 12.1, 14.1)
player <- seq_along(pct_body)
df <- data.frame(player = player, pct_body = pct_body)
df
```


## Calculate z-score

```{r}
m <- mean(df$pct_body)
s <- sd(df$pct_body)

df.calc <- df %>%
mutate(zscore = round((pct_body - m) / s, 2))
df.calc
```

Notice that you end up with negative z-scores and positive z-scores. Negative z-scores represent values below the mean, and positive z-scores represent values above the mean. 


The grades for a statistics exam are as follows: 3, 5, 5, 6, 7.5, 6, 5, 1, 10, 4. Which score is an outlier? (Use the interquartile range (IQR).)

```{r}
grades <- c(3, 5, 5, 6, 7.5, 6, 5, 1, 10, 4)
```


```{r}
iqr = IQR(grades)
iqr
quan <- quantile(grades)
quan
```

```{r}
upper <- quantile(grades)[4] + 1.5 * iqr
upper

lower <- quantile(grades)[2] - 1.5 * iqr
lower
```

```{r}
boxplot(grades)
```

How many goals have the top strikers of the Dutch Eredivisie football competition scored? We look at 10 strikers and obtained the following information: 12, 10, 11, 12, 11, 14, 15, 18, 21, 11. The (1) ... of the dataset equals 12, the mean equals (2) ... and the (3) ... equals 11. The standard deviation equals (4) ... Fill in the right words/numbers on the dots.


```{r}
goals <- c(12, 10, 11, 12, 11, 14, 15, 18, 21, 11)
```


```{r}
mean(goals)
sd(goals)
IQR(goals)
quantile(goals)
```

```{r}
# Create the function for the mode
getmode <- function(v) {
   uniqv <- unique(v)
   uniqv[which.max(tabulate(match(v, uniqv)))]
}
```


```{r}
getmode(goals)
```

