---
title: "R Notebook"
output: html_notebook
---


```{r}
library(prob)  #load the package
```



# 1. Sample Spaces

```{r}
#--------------------------------------------------------
#   Probability
#--------------------------------------------------------
#Consider the random experiment of tossing a coin. The outcomes are H and T. We can set
#up the sample space quickly with the tosscoin function:

tosscoin(1) 

tosscoin(3)

rolldie(1)
cards()
```

## Sampling from Urns
Let our urn simply contain three balls, labeled 1, 2, and 3, respectively. We are
going to take a sample of size 2 from the urn

```{r}
urnsamples(1:3, size = 2, replace = TRUE, ordered = TRUE)

#This experiment is equivalent to rolling a 3-sided die twice, which we could have accomplished with
rolldie(2, nsides = 3)
```

# 2. Events

```{r}
S <- tosscoin(2, makespace = TRUE)
S[1:3, ] 
S[c(2,4),]

```


```{r}
S<-cards()
subset(S,suit=="Heart")


```

```{r}
subset(S, rank %in% 7:9)

subset(rolldie(3), X1 + X2 + X3 > 16)



```

```{r}
A = subset(S, suit == "Heart")
B = subset(S, rank %in% 7:9)
union(A,B)


```

```{r}
intersect(A,B)

```

```{r}
setdiff(B,A)

setdiff(S,A)
```

# 3. Computing probability

## The Birthday Problem
Suppose that there are `n` people together in a room.
Each person announces the date of his/her birthday in turn. 
The question is: what is the probability of at least one match? 

```{r}
g <- sapply(1:50, pbirthday)
plot (1:50, g,
      xlab = "Number of people in room ",
      ylab = "Prob (at least one match )",
      main = "The Birthday Problem")
abline (h = 0.5)
abline (v = 23, lty = 2) # dashed line
```

## conditional probability
```{r}
S <- rolldie(2, makespace = TRUE) # assumes ELM
head(S)

```

```{r}
A <- subset(S, X1 == X2)      # die1 == die2
B <- subset(S, X1 + X2 >= 8)  # di1+die2 >= 8

Prob(A, given = B)  # probability of two dice the same given >= 8
Prob(S, X1==X2, given = (X1+X2>=8))
```



